MLH1 Results Outline

### 2020-02-26 doc/Results.rmd
small changes and notes

@@

<!-- which maybe be attributed to noise effects 
The ratios for MLH1:DMC1^Leptotene^ and MLH1:DMC1^zygotene^ are shown in Table X, and can be used as a proxy for the CO:NCO ratio (instead CO:(CO+NCO) ratio). -->

@@

<come up with a better section title>

@@

**M3. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}

@@

```{r Q1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

(rethink where this section should go)
new outline

1. illustrate problem(affects mostly SC length)

2. Expected impact on sex comparisons, estimated effect size of the X  
3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons

4. 19 female, 19male, 20female 20 male

The female mouse averages should have adjustments for the XX. working on code to estimate the SC length from 3rd largest bivalent from female whole cell data across strains. Subtract this amount from the female mouse averages ... This isn't the best solution -- since I can't determine what proportion of cells for female mouse averages include the XX, (most cells are missing at least 3 bivalents)

- Of all female single bivalents observations, 5% are XX (1 of 20).

- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).

- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.

I think a formula something like this can be applied to adjust for XX

- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX /*

@@

<transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents
1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
-- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).

>>>>

after DSB formation, DSBs are repaired as either NCO or CO (CO designation step)
The vast majority of DSBs are repaired as NCOs, with a minority (in most species) of approximately 10% are repaired as crossovers.

-DSBs are either repaired as crossovers or noncrossovers (NCO), with conversion tract lengths (30 to 100bp long). (this is also known as the CO repair deicison )

thus the ratio of CO:DSB is in part an indicator of the proportion of DSBs which are designated into COs -- (and the others are repaired as NCO).

@@

@@

@@

**M1. Mixed Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex + rand(strain) + \varepsilon $$

**M2.Linear Model** $$mouse \ average \ SC \ metric ~=~ subsp * sex * strain + \varepsilon $$

```{r Q1.M.lmer_totalSC, echo=FALSE, warning=FALSE, message=FALSE}

@@

```{r Q1.M1_short_biv_lmer, echo=FALSE, warning=FALSE, message=FALSE}
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

@@

1. illustrate problem(affects mostly SC length)  
2. Expected impact on sex comparisons, estimated effect size of the X  
3. (prove general pattern that ALL bivalents are longer), chrms sorted by bin comparisons
4. permutations of 19 female, 19male, 20female 20 male

<
- Of all female single bivalents observations, 5% are XX (1 of 20). 
- The XX is large, likely within the top 25% longest bivalents of the cell (3rd largest by Mb).
- The average % of XX for whole cell SC (sum(all bivalents)) can be calculated from the whole.cell data set. Lets guess 12% of a cell's total SC area is XX.
- rate of bivalent segmentation /* rate of XX, 5% /* mean SC length for 3rd longest bivalent / total SC area (by bivalent) = proportion of SC area due to XX>

@@

<transition>  -- following up on SC area - with how COs are place --- placement and number of COs on single bivalents

1. Transition, --- the SC-AE area can be thought of as the 'area / available real-estate for COs
-- these results suggest that this area is 'utilized' in different ways across sexes and strains - so we next investigate the recombination landscape - (ie the relationship between the placement and number of COs along chromosomes).
